---
output:
  html_document: default
  word_document: default
---
##1. Code for reading in the dataset and/or processing the data
      Add Libraries and set up working Directory, download the source/data              file,unzip the file, read the data file, look at what data looks like

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(ggplot2)
path<- file.path(getwd(),"activity_monitoring.zip")
url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, path, method="libcurl")
unzip(path)
dir()
data<- read.csv("activity.csv",header=TRUE,sep=',')
head(data)
str(data)
summary(data)

```
The above summary indicates 2304 NA's in 17568 observations in steps variable, with Mean for Steps as 37.38. The presence of NA's has made the median value of steps equal to Zero



 Tidying data of daily steps 
```{r}
library(lubridate)
data$steps <- as.numeric(data$steps)
data$date<- as.character(data$date)
data$date <- ymd(data$date)
data$interval<-as.numeric(data$interval)
sum_steps_daily <- data%>%group_by(date)%>% summarise(sumsteps=sum(steps,na.rm=TRUE))
head(sum_steps_daily)
```

## 2.Histogram of the total number of steps taken each day
    Use hist function to Plot histogram of Steps 

```{r}
with(sum_steps_daily,hist(sumsteps,col= "Blue", xlab= "daily steps", main="Histogram of daily steps"))

```

##3. To find mean and median of the steps taken per day
```{r}
steps_mean <- mean(sum_steps_daily$sumsteps,na.rm=TRUE)
steps_median <-median(sum_steps_daily$sumsteps,na.rm=TRUE)
print( steps_mean)
print(steps_median)
```
##4. Daily average activity pattern
```{r}
library(ggplot2)
fivemts_averages <- aggregate(x = list(steps = data$steps), by = list(interval = data$interval), 
    FUN = mean, na.rm = TRUE)
ggplot(data = fivemts_averages, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute interval") + 
    ylab("average number of steps taken")
```
##5. To find Max activity of steps in a 5minute interval
```{r}
fivemts_averages[which.max(fivemts_averages$steps),]
```
##6.Imputing missing values with daily avearge in 5 minute inetrval
As indicated through summary of data there are 2304 NAs in variable steps
```{r}
nodata<-is.na(data$steps)
int_average<-tapply(data$step,data$interval,FUN=mean,na.rm=TRUE)
data_wnas<- data
data_wnas$steps[nodata]<-int_average[as.character(data_wnas$interval[nodata])]
summary(data_wnas)
```
##7. To compute average steps and histogram after removing NAs

```{r}
steps_wnas <- data_wnas %>%
group_by(date) %>%
summarize(sum_steps = sum(steps)) %>%
print
with(steps_wnas,hist(sum_steps,col= "violet", xlab= "daily steps", main="Histogram of daily steps after replacing NAs"))
```
compute mean and median values

```{r}
mean_steps_wnas <-mean(steps_wnas$sum_steps,na.rm=TRUE)
mean_steps_wnas
median_steps_wnas <- median(steps_wnas$sum_steps,na.rm=TRUE)
median_steps_wnas
```
The Values of mean(9354 vs 10766) and Median(10395 vs 10766) have significantly increased after replacing NAs with average value of steps in that time interval ex. 0,5,10...
##8. Panel plot for steps on weekdays,weekends

using function weekdays and data with replaced NA's
```{r}
data_wnas$date <- as.Date(data_wnas$date)
data_daytype<- with(data_wnas,mutate(data_wnas,Daytype=ifelse(weekdays(data_wnas$date) =="Saturday"| weekdays(data_wnas$date)=="Sunday" , "Weekend", "Weekday")))
data_daytype$Daytype <- as.factor(data_daytype$Daytype)
steps_daytype <- data_daytype %>%
group_by(Daytype,interval) %>%
summarize(avg_steps = mean(steps)) 
str(steps_daytype)
g<-ggplot(steps_daytype, aes(x=interval, y=avg_steps, color = Daytype)) +
  geom_line() +
  facet_wrap(~Daytype, ncol = 2, nrow=1)
print(g)
```